Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-10-01 arxiv/proof_examples.tex
revise

 &=\tr\left[  \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1} \left( \hat\Sigma- A^2 (X^{(1)})^\top X^{(1)} \right)^2\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}\left( \hat\Sigma- A^2 (X^{(1)})^\top X^{(1)} \right) \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] \\
 &=\tr\left[  (X^{(1)})^\top X^{(1)}\right] - A^2\tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)}\right] = \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right] .

@@

Using the concentration estimates in Lemma \ref{largedeviation}, we can control the difference between $g(A)$ and its partial expectation. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(A) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
$$\left| g(A) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2 }\right| \le p^{-c_\infty}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right). $$
 \end{claim}
 \begin{proof}
 \todo{add proof}
 \end{proof}

@@

\begin{proof}
Corresponding to \eqref{EgA}, we define the function
\begin{align*}
h(A):= [(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right].
\end{align*}
Using Fact \ref{lem_minv} (iii), we can obtain that with high probability,
\begin{align*} 
&\left( A^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}\\
&\preceq (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \\
&\preceq \left( A^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}.
\end{align*}
Then we get that
\be\label{eq_h1}h(1)= d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2\left(\rho_1^{-1} + \rho_2^{-1}\right)^{-1}\lesssim pd^2.\ee
where we used that $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$. On the other hand, if $|A-1|\ge \delta$ for a small constant $\delta>0$, then we have 
\begin{align}\label{eq_hA}
h(A)\gtrsim (1+A^2)\kappa^2 \cdot p\left(\rho_1^{-1} + A^2 \rho_2^{-1}\right)^{-1} \gtrsim p\kappa^2 ,
\end{align}
where in the first step we used that $|A-1|^2\gtrsim 1+A^2$ for $|A-1|\ge \delta$, $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$, and in the second step we used \eqref{choiceofpara0}. Hence using Claim \ref{claim_largedev1}, \eqref{eq_h1} and \eqref{eq_hA}, we obtain that with high probability,
\begin{align*}
g(A)&\ge  h(A) +\sigma^2(n_1+n_2-p) -p^{-c_\infty}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \\
&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-c_\infty}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1).
\end{align*}
This gives that
\be\label{rough A-1}
|\hat A-1|\le \delta
\ee
for any small constant $\delta>0$.

{\cor require $k^2\gg d^2 + p^{-1/2+c}\sigma^2$}

Hence it remains to consider the region $|A-1|\le \delta$ for a small enough constant $\delta>0$.

We first find the minimizer of $h(A)$, denoted by $A^\star$. 

@@

@@

@@

Next we prove \eqref{hatw_add1}.

Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%\begin{lemma}
%For the isotropic model, we have
\be\label{hatv_add1}
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
\ee
%\end{lemma}
%\begin{proof}
 In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under \eqref{choiceofpara} we have
$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
Thus we obtain that for $|w-1|\ge 2C\cal E$,
$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
which leads to $g(w) > g(\hat w)$ with high probability by \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with \eqref{hatw_add1}, we conclude \eqref{hatv_add1}.
%\end{proof}
\end{proof}

>>>>

 &=\tr\left[  (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} \right] \\
 & = \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right] .

@@

 In the first step of the above derivation, we used that 
 $$(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} = (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)}.$$

Using the concentration estimates in Lemma \ref{largedeviation}, we can control the difference between $g(A)$ and its partial expectation. %$\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{g(A) \mid X_1, X_2 ,\beta_1,\beta_2}$.
 \begin{claim}\label{claim_largedev1}
 In the setting of Proposition \ref{cor_MTL_loss}, we have that with high probability,
$$\left| g(A) - \exarg{\epsilon^{(1)}, \epsilon^{(2)} ,\beta_1,\beta_2}{g(A) \mid X_1, X_2 }\right| \le p^{-1/2+c}\cdot p\left(\sigma^2 + \kappa^2+d^2 \right) $$
for any small constant $c>0$.
 \end{claim}
 \begin{proof}
 \todo{add proof}
 \end{proof}

@@

\begin{proof}
Corresponding to \eqref{EgA}, we define the function
\begin{align*}
h(A):= [(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right].
\end{align*}
Using Fact \ref{lem_minv} (iii), we can obtain that with high probability,
\begin{align} 
&\left( A^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1} \nonumber\\
&\preceq (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \nonumber\\
&\preceq \left( A^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}. \label{op_uplow}
\end{align}
Then we get that
\be\label{eq_h1}h(1)= d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2\left(\rho_1^{-1} + \rho_2^{-1}\right)^{-1}\lesssim pd^2.\ee
where we used that $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$. On the other hand, if $|A-1|\ge \delta$ for a small constant $\delta>0$, then we have 
\begin{align}\label{eq_hA}
h(A)\gtrsim (1+A^2)\kappa^2 \cdot p\left(\rho_1^{-1} + A^2 \rho_2^{-1}\right)^{-1} \gtrsim p\kappa^2 ,
\end{align}
where in the first step we used that $|A-1|^2\gtrsim 1+A^2$ for $|A-1|\ge \delta$, $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$, and in the second step we used \eqref{choiceofpara0}. Hence using Claim \ref{claim_largedev1}, \eqref{eq_h1} and \eqref{eq_hA}, we obtain that with high probability,
\begin{align*}
g(A)&\ge  h(A) +\sigma^2(n_1+n_2-p) -p^{-1/2+c}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \\
&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-1/2+c}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1)
\end{align*}
under \eqref{choiceofpara0}.
This gives that
\be\label{rough A-1}
|\hat A-1|\le \delta
\ee
for any small constant $\delta>0$.

%Hence it remains to consider the region $|A-1|\le \delta$ for a small enough constant $\delta>0$. 
To obtain the better bound \eqref{hatw_add1}, we study the function $h(A)$ more closely and find its minimizer of $h(A)$, denoted by $A^\star$. 

@@

\begin{align}\label{h'(A)1}

@@

\begin{align}\label{h'(A)2}

@@

Now we are ready to prove \eqref{hatw_add1}. We choose the constant $c$ in Claim \eqref{claim_largedev1} small enough such that $1/2-c> c_\infty$. First using \eqref{op_uplow} and \eqref{h'(A)1}, we obtain that for $1+d^2/\kappa^2 \le A\le 1+\delta$,
$$h'(A)\ge \frac{2(A-1)\kappa^2 - d^2}{A^3} \cdot p^{-1} \tr\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right) \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Thus we get that if $1+d^2/\kappa^2 +p^{-c_\infty}\le A\le 1+\delta$,

Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%\begin{lemma}
%For the isotropic model, we have
\be\label{hatv_add1}
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
\ee
%\end{lemma}
%\begin{proof}
 In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under \eqref{choiceofpara} we have
$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
Thus we obtain that for $|w-1|\ge 2C\cal E$,
$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
which leads to $g(w) > g(\hat w)$ with high probability by \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with \eqref{hatw_add1}, we conclude \eqref{hatv_add1}.
%\end{proof}

{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}
